---
title: "Progress Report - <simeonbundt14>"
author: "Simeon Bundt"
date: "12-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(Lahman)
```
# Progress Report - <simeonbundt14>

## Objectives
- Scrape Hall of Fame data for 2024 using `rvest`.
- Clean the data to match the `HallOfFame` table in the `Lahman` package.
- Save the cleaned data as a CSV file.

## Steps Taken

### 1. Load Required Libraries
```{r}
url <- "https://www.baseball-reference.com/awards/hof_2024.shtml"
html <- read_html(url)
tables <- html_table(html)
data <- tables[[1]]
head(data)


# Extract column names from the first row and make them unique
colnames(data) <- make.unique(as.character(unlist(data[1, ])))  
data <- data[-1, ]

# Rename columns to match the HallOfFame structure in Lahman
cleaned_data <- data %>%
  rename(
    playerID = Name,
    yearID = `YoB`,
    votes = Votes,
    inducted = `%vote`
  ) %>%
  mutate(
    yearID = 2024,
    votes = as.numeric(votes),
    inducted = if_else(as.numeric(str_remove(inducted, "%")) >= 75, "Y", "N"),
    category = "Player",
    votedBy = "BBWAA"
  )
# Check the cleaned data
head(cleaned_data)

# Combine the new data with existing HallOfFame data
updated_HallOfFame <- bind_rows(HallOfFame, cleaned_data)

# Preview the combined data
head(updated_HallOfFame)

# Save the cleaned and combined data as a CSV file
write_csv(updated_HallOfFame, "HallOfFame.csv")

# Create a bar plot of Hall of Fame inductions over the years
updated_HallOfFame %>%
  ggplot(aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(1936, 2024) +
  labs(
    title = "Hall of Fame Inductions Over the Years",
    x = "Year",
    y = "Number of Inductions",
    fill = "Inducted"
  )
```

### Challenges Faced
- Parsing the HTML tables took time due to inconsistent formatting.
- Some columns had unexpected data types, which caused errors in data processing.
- The scraped table lacked default values for `category` and `votedBy`.
- Duplicated Column Names: While trying to clean the scraped data, duplicate column names caused an error in the rename() function.
- Non-Character Column Names: An attempt to use make.unique() on the first row to resolve the duplicated column names resulted in another error because make.unique() expects a character vector, but the first row was a data frame row.


### Resolutions
- Adjusted column names and added missing fields manually.
- Used `mutate` to clean and process variables for consistency.
- Verified the final data structure matches the `HallOfFame` table.
- Used make.unique() to ensure all column names were unique before renaming them.
- After making column names unique, proceeded with renaming them to match the Lahman HallOfFame dataset structure using rename().
